---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


# reproducible research: week 2 - activity 11

## Activate libraries
```{r}
library(tidyverse)
library(lubridate)
library(missForest)
library(ggplot2)
```

## Read in the data
```{r}
dta <- read.table(file = "activity.csv", header = TRUE, sep = ",")

dta$date <- as.Date(dta$date)

dta$month <- month(dta$date)
dta$year  <- year(dta$date)
dta$day   <- day(dta$date)
```

## Do a histogram of the steps each day
```{r}
dta %>% 
  group_by(date) %>%
  summarise(total_steps = sum(steps, na.rm = TRUE)) ->
  daily_steps

hist(daily_steps$total_steps)
```

## Mean and median of the steps each day
```{r}
daily_steps %>%
  summarise(steps_mean = mean(total_steps, na.rm = TRUE), 
            steps_median = median(total_steps, na.rm = TRUE)) ->
  steps_mean_median

print(steps_mean_median)
```


## Time series plot of the average number of steps taken
```{r}
dta %>% 
  group_by(date) %>%
  summarise(steps_mean = mean(steps, na.rm = TRUE)) ->
  dta_mean_daily

plot(y = dta_mean_daily$steps_mean, x = dta_mean_daily$date, type = "l")
```



## The 5-minute interval that, on average, contains the maximum number of steps

```{r}
  dta %>%
    group_by(interval) %>%
    summarise(steps_mean = mean(steps, na.rm = TRUE)) ->
    steps_mean_interval
    
steps_mean_interval[which.max(steps_mean_interval$steps_mean), "interval"]
```

## Code to describe and show a strategy for imputing missing data

There are several possibilities to impute missing data. Contrary to normal regression when imputing data normally an 'elephant' model is used, meaning also using variables in the regression model to describe missing variables which are normally not used, e.g. due to low p-value.

In terms of packages there are many packages out there to impute missing data. One simple solution is the missForest package which uses random forests to impute missing data. Other possibilities could be the mice package or the midas package.

Generally the rule holds that when building after the imputation a model, at least all the variables which will be used in the subsequent model should be used in the imputation process/model.
```{r}
dta_matrix <- as.matrix(dta %>% select(-one_of("date")))

dta_imputed <- missForest(dta_matrix)
dta_imputed <- as.data.frame(dta_imputed$ximp)
dta_imputed$date <- as.Date(paste0(dta_imputed$year, "-", dta_imputed$month, "-", dta_imputed$day))

```

## Histogram of the total number of steps taken each day after missing values are imputed

```{r}
dta_imputed %>% 
  group_by(date) %>%
  summarise(total_steps = sum(steps, na.rm = TRUE)) ->
  daily_steps_imputed

hist(daily_steps_imputed$total_steps)
```


## Make a panel plot containing a time series plot after creating weekday/weekend

```{r}
dta_wd <- dta_imputed %>%
  mutate(wd = ifelse(weekdays(date) %in% c("Samstag", "Sonntag", "Sunday", "Saturday"), 
                     "weekend", "weekday"))

dta_wd <- aggregate(steps ~ interval + wd, data = dta_wd, mean)
ggplot(dta_wd, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(wd ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")
```


